Quantifying the importance of an inference model in Bayesian phylogenetics

### 2019-12-01 content.tex
Verify all code works

\begin{sidewaystable}
\centering
  \begin{tabular}{|p{3.4cm}|p{9.7cm}|p{4.5cm}@{}|}
    \hline
    \centering
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Sub-argument} & 
    \textbf{Description} &
    \textbf{Possible values} \\ 
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;tree_prior; &
    Macroevolutionary diversification model &
    Yule, BD, CBS, CCP, CEP \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;clock_model; &
    Clock for the DNA mutation rates &
    strict, RLN \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;site_model; &
    Nucleotide substitution model &
    JC, HKY, TN93, GTR \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;mutation_rate; &
    Pace at which mutation occurs &
    \verb;mutation_rate; $\in \mathbb{R}_{>0}$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;root_sequence; &
    DNA sequence at the root of the tree &
    any combination of a, c, g, t \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;model_type; &
    Criterion to select an inference model &
    Generative, Candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;run_if; &
    Condition under which an inference model is used &
    Always, Best candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;do_measure_evidence; &
    Sets whether or not the evidence of the model is to be computed &
    TRUE, FALSE \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;error_function; &
    Specifies how to measure the error &
    nLTT, $|\gamma|$ \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;burn_in_fraction; &
    Specifies the percentage of initial posterior trees to discard &
    \verb;burn_in_fraction; $\in [0, 1]$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}
  \caption{
    Most important parameter options.
    Yule = pure birth model (\cite{yule})
    BD = birth death (\cite{nee1994reconstructed}), 
    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
    CCP = coalescent constant-population, 
    CEP = coalescent exponential-population,
    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
    TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
    GTR = Generalized time-reversible model (\cite{tavare1986some})
    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}).
  }
  \label{tab:options}
\bigskip

@@

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single
]
usethis::install_github("richelbilderbeek/pirouette")
\end{lstlisting}
To start using \verb;pirouette;, 
load its functions in the global namespace first:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
library(pirouette)
\end{lstlisting}
Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
This can be done from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install_beast2()
\end{lstlisting}
For the option to select the best candidate model,
\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
It can be installed from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install_beast2_pkg("NS")
\end{lstlisting}

@@

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single,
    label = {lst:create_alignment_params}, 
    caption = {Create an alignment.}
  ]
alignment_params <- create_alignment_params(
  sim_true_alignment_fun = get_sim_true_alignment_with_std_site_model_fun(
    mutation_rate = 0.1,
    site_model = create_jc69_site_model()
  ),
  root_sequence = create_blocked_dna(length = 1000),
  rng_seed = rng_seed
)
\end{lstlisting}

@@

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_gen_experiment_explicit},
  caption = {
    Create an experiment with the generative model,
    that will always be used in the actual inference, 
    using explicit arguments.
  }
]
generative_experiment <- create_experiment(
  inference_conditions = create_inference_conditions(
    model_type = "generative", 
    run_if = "always"
  ), 
  inference_model = create_inference_model(
    tree_prior = create_yule_tree_prior(),
    clock_model = create_strict_clock_model(), 
    site_model = create_jc69_site_model(),
    mcmc = create_mcmc(store_every = 1000)
  )
)
\end{lstlisting}

@@

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_error_measure_params},
  caption = Calling \texttt{create\_error\_measure\_params}.
]
error_measure_params <- create_error_measure_params(
  error_function = get_nltt_error_function(),
  burn_in_fraction = 0.1
)
\end{lstlisting}

@@

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_twinning_params},
  caption = Create the default twinning parameters.
]
twinning_params <- create_twinning_params(
  twin_model = "birth_death",
  method = "random_tree"
)
\end{lstlisting}

>>>>

\begin{sidewaystable}
\centering
  \begin{tabular}{|p{3.4cm}|p{9.7cm}|p{4.5cm}@{}|}
    \hline
    \centering
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Sub-argument} & 
    \textbf{Description} &
    \textbf{Possible values} \\ 
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;tree_prior; &
    Macroevolutionary diversification model &
    Yule, BD, CBS, CCP, CEP \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;clock_model; &
    Clock for the DNA mutation rates &
    strict, RLN \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;site_model; &
    Nucleotide substitution model &
    JC, HKY, TN93, GTR \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;mutation_rate; &
    Pace at which mutation occurs &
    \verb;mutation_rate; $\in \mathbb{R}_{>0}$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;root_sequence; &
    DNA sequence at the root of the tree &
    any combination of a, c, g, t \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;model_type; &
    Criterion to select an inference model &
    Generative, Candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;run_if; &
    Condition under which an inference model is used &
    Always, Best candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;do_measure_evidence; &
    Sets whether or not the evidence of the model is to be computed &
    TRUE, FALSE \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;error_fun; &
    Specifies how to measure the error &
    nLTT, $|\gamma|$ \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \verb;burn_in_fraction; &
    Specifies the percentage of initial posterior trees to discard &
    \verb;burn_in_fraction; $\in [0, 1]$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}
  \caption{
    Most important parameter options.
    Yule = pure birth model (\cite{yule})
    BD = birth death (\cite{nee1994reconstructed}), 
    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
    CCP = coalescent constant-population, 
    CEP = coalescent exponential-population,
    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
    TN93 = Tamura and Nei (\cite{tamura1993estimation}), 
    GTR = Generalized time-reversible model (\cite{tavare1986some})
    RLN = relaxed log-normal clock model (\cite{drummond2006relaxed}).
  }
  \label{tab:options}
\bigskip

@@

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single
]
remotes::install_github("richelbilderbeek/pirouette")
\end{lstlisting}
To start using \verb;pirouette;, 
load its functions in the global namespace first:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
library(pirouette)
\end{lstlisting}
Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
This can be done from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install_beast2()
\end{lstlisting}
For the option to select the best candidate model,
\verb;pirouette; needs the "NS" BEAST2 package [\cite{maturana2018model}].
It can be installed from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
install_beast2_pkg("NS")
\end{lstlisting}

@@

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single,
    label = {lst:create_alignment_params}, 
    caption = {Create an alignment.}
  ]
alignment_params <- create_alignment_params(
  sim_true_alignment_fun = get_sim_true_alignment_with_std_site_model_fun(
    mutation_rate = 0.1,
    site_model = create_jc69_site_model()
  ),
  root_sequence = create_blocked_dna(length = 1000)
)
\end{lstlisting}

@@

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_gen_experiment_explicit},
  caption = {
    Create an experiment with the generative model,
    that will always be used in the actual inference, 
    using explicit arguments.
  }
]
generative_experiment <- create_experiment(
  inference_conditions = create_inference_conditions(
    model_type = "generative", 
    run_if = "always"
  ), 
  inference_model = create_inference_model(
    tree_prior = create_yule_tree_prior(),
    clock_model = create_strict_clock_model(), 
    site_model = create_jc69_site_model(),
    mcmc = create_mcmc()
  )
)
\end{lstlisting}

@@

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_error_measure_params},
  caption = Calling \texttt{create\_error\_measure\_params}.
]
error_measure_params <- create_error_measure_params(
  error_fun = get_nltt_error_fun(),
  burn_in_fraction = 0.1
)
\end{lstlisting}

@@

\begin{lstlisting}[
  language = R,
  floatplacement = ht,
  frame = single,
  label = {lst:create_twinning_params},
  caption = Create the default twinning parameters.
]
twinning_params <- create_twinning_params(
  sim_twin_tree_fun = get_sim_bd_twin_tree_fun(),
  sim_twin_alignment_fun = get_sim_twin_alignment_with_std_site_model_fun()
)
\end{lstlisting}
